---
output: github_document 
author: Charisa Geyer (22894551)
date: 16-06-2025
---

# Purpose

This is the main README for the mock exam where I will outline the process of data wrangling and obtaining all results. Each question is saved as its own project, with its own texevier/ readme files. 


```{r, echo=FALSE}
rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(tidyverse)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
```
# Create Question folders using Texevier
```{r}

Texevier::create_template("C:/Users/Charisa/OneDrive - Stellenbosch University/Documents/2025/Semester 1/Data_Science/22894551_Mock", "Question 1")

Texevier::create_template("C:/Users/Charisa/OneDrive - Stellenbosch University/Documents/2025/Semester 1/Data_Science/22894551_Mock", "Question 2")

Texevier::create_template("C:/Users/Charisa/OneDrive - Stellenbosch University/Documents/2025/Semester 1/Data_Science/22894551_Mock", "Question 3")

```

# Question 1: World Happiness Report


I put the unzipped data files in their respective question folders.

I must collate the data into a single dataframe. 
I created a function called data_collating to do so.
Remember to run the list.files thing again to load the function!

```{r}
Datroot <- "C:/Users/Charisa/OneDrive - Stellenbosch University/Documents/2025/Semester 1/Data_Science/22894551_Mock/Question 1/data/Happy"
datcolat <- data_collating(Datroot)
```

## (a) First plot
We want to plot per region, the Ladder Score, upperwhisker and lowerwhiskers using ggplot. Also,
add directly above each region’s plot the average Healthy Life Expectancy. Your plot
should also be arranged by Average Life Expectancy.
Let's create a function to create the errorgeombar. 
```{r, warning =  FALSE, fig.align = 'center', fig.cap = "Happiness Ladder.\\label{Figure1}", fig.ext = 'png', fig.height = 5, fig.width = 7}

q1 <- geom_errorbar_plot(datcolat, xaxis_size = 5, xaxis_rows = 3)

q1

```

## (b) Second plot
Create a barplot that shows the breakdown of Ladder scores per region. Arrange the
regions as they appear in the plot by Ladder score. Also, add South Africa’s ladder to
this plot (make SA the first bar).
First I created a function that puts the data into a workable format with all components specified and only the relevant columns selected. 
Then I create the graph using ggplot. 


```{r}
q1b <- Barplot_breakdown(datcolat, Title = "Breakdown of WHR Ladder Score by region", Subtitle ="Comparison with South Africa", xaxis_size = 9, xaxis_rows = 3)

q1b
```

<!-- It needs to show what contributes to the Ladder score. so it is best to do a stacked bargraph. -->
<!-- ```{r} -->

<!-- ``` -->


# Question 2 - Wine Whine Wine

Task: write a short piece on the preferences
of Sommeliers globally, and also show which wines and regions are preferred in South Africa

Plot how many ratings each country received as a barplot, with the median score placed
vertically above each bar.
```{r}
q2 <- reviewcount_barplot(WineMag, xaxis_rows = 1)
q2
```

